\documentclass[]{report}

\voffset=-1.5cm
\oddsidemargin=0.0cm
\textwidth = 480pt

\usepackage{framed}
\usepackage{subfiles}
\usepackage{graphics}
\usepackage{newlfont}
\usepackage{eurosym}
\usepackage{amsmath,amsthm,amsfonts}
\usepackage{amsmath}
\usepackage{color}
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\begin{document}
%-------------------------------------------------------- %
\begin{verbatim}
5 Introduction to Graph Theory 
5.1 What is a graph?
5.1.1 Some definitions  
5.1.2 Degree of a vertex 
5.1.3 Some special graphs  

5.2 Paths, cycles and connectivity  
5.2.1 Paths 
5.2.2 Cycles  
5.2.3 Connectivity  

5.3 Isomorphism of graphs  
5.3.1 Showing that two graphs are isomorphic 
5.3.2 Showing that two graphs are not isomorphic
 
5.4 Adjacency matrices and adjacency lists  
5.4.1 Adjacency matrix of a graph 
5.4.2 Adjacency lists 

\end{verbatim}

\section*{What is Graph Theory}
Graph theory is the study of graphs, which are mathematical structures used to model pairwise relations between objects. A "graph" in this context is made up of "vertices" or "nodes" and lines called edges that connect them. A graph may be undirected, meaning that there is no distinction between the two vertices associated with each edge, or its edges may be directed from one vertex to another

%-----------------------------------------------------%
\section*{Session 05:Graphs}
\begin{itemize}
\item[5A.1] What is a Graph?
\item[5A.2] Paths Cycles and Connectivity
\item[5A.3] Isomorphisms of a graph
\item[5A.4] Adjacency Matrices and Adjacency Lists
\end{itemize}



\subsection*{Isomorphism}
\begin{itemize}
\item They have a different number of connected components
\item They have a different number of vertices
\item They have different degrees sequences
\item They have a different number of paths of any given length
\item They have a different number of cycles of any length.
\end{itemize}

\begin{itemize}
\item Spanning Subgraphs of G.
\item a vertex is said to be an \textbf{emph{ isolated vertex}} if it has a degree of zero.
\item a vertex is said to be an \textbf{emph{ end-vertex}} if it has a degree of one.
\item a vertex is said to be an \textbf{emph{ even vertex}} if it has a degree of an even number.
\item a vertex is said to be an \textbf{emph{ odd vertex}} if it has a degree of an odd number.
\item A graph is said to be \textbf{emph{k-regular}} if the degree of each vertex is $k$. 
\item Every Graph has an even number of odd vertices.
\item A cubic graph is a graph where every vertex has degree three.
\end{itemize}


\end{document}
